---
title: "DSO Final Report - MobileOS"
author: "Michelle Offir Geveye"
date: "December 4, 2016"
output: pdf_document
---

# Mobile Channels Analysis

```{r}
myLA311 = read.csv("MyLA311_Service_Request_Data_2016.csv")

library(dplyr)
library(lubridate)
library(ggplot2)

## Filter data with only Mobile App Requests
data = filter(myLA311, MobileOS != "")


```

## Check Mobile requests bw Sep 15 and Sep 16

```{r}

data$CreatedDate = mdy_hms(data$CreatedDate)

MobileYear = filter(data, CreatedDate >= "2015-09-01 00:00:00" & CreatedDate < "2016-08-31 00:00:00")

ggplot(MobileYear, aes(x = CreatedDate, fill = MobileOS)) + 
  geom_density( alpha = 0.4, color = NA) + 
  geom_line(stat = "density", size = 0.5) +
  ggtitle("Mobile Requests Created From September 2015 to September 2016") +
  xlab("Date") 
```


Seems like iOS was not existing before Jan 2016.

## Check most common requests

```{r}
Reqs = data %>%
  group_by(RequestType, MobileOS) %>%
  summarise(count = n())

ggplot(Reqs, aes(x = RequestType, y = count, fill = MobileOS)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  ggtitle("Most Comon Requests by Mobile OS") +
  xlab("Total Requests") +
  ylab("Request Type")
```

  
Most requested: Bulky Items (Android), Graffiti Removal (iOS), Illegal Dumping Pickup (Android), Other (iOS)

## Check requests by day and time

```{r}

data$dow = wday(data$CreatedDate, label = T)
data$crHr = hour(data$CreatedDate)
heat = data %>%
  group_by(dow, crHr, MobileOS) %>%
  summarise(count = n())

ggplot(heat, aes(x = dow, y = factor(crHr), fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "darkred") +
  ggtitle("Distribution of Requests by Day and Time") +
  xlab("Day of Week") +
  ylab("Hour of Day") +
  facet_grid(~MobileOS)
```
Most requests with Android, Monday is a very busy day. Most calls bw 8AM and 3pm. iOS calls go through 9 PM, Barely and calls from midnight to 6 AM.

## Filter night calls

```{r}
data$crHr = hour(data$CreatedDate)
night = filter(data, crHr >=0 & crHr<6)


night = night %>%
  filter(crHr >=0 & crHr<6) %>%
  group_by(RequestType, crHr, MobileOS) %>%
  summarise(count = n())

ggplot(night, aes(x = crHr, y = RequestType, fill = count)) + 
  geom_tile() + 
  scale_fill_gradient(low = "white", high = "darkred")+
  facet_grid(~MobileOS) +
  ggtitle("Common Requests Made Between Midnight and 6 AM") +
  ylab("Request Type") +
  xlab("Hour Created")
```

Most common request at night - Bulky Items followed by Graffiti Removal. More requests made from Android at night.


## Check Request Source

```{r}
reqSource = data %>%
  group_by(MobileOS, RequestSource) %>%
  summarise(count = n()) %>%
  mutate(pct=100*count/sum(count))

reqSource
```

99.9% of requests are made via the mobile app. The other sources are not enough to explore

## Check Mobile Map usage in LA

```{r}

library(ggmap)

LA = "Los Angeles"
laMap = qmap(LA, color = "bw")

laMap +
  geom_point(data = data,
             aes(Longitude, Latitude, color = MobileOS),
             size = 1, alpha = 0.1) +
  ggtitle("Mobile Request Usage Accross LA")


```
iOS seems to be used more on the map (which is weid because in the graphs it looks like the other way around). Mobile use failry spread. Concentration in Center LA area, SM, and Vally.


## Check daily usage by MobileOS in LA

```{r}

laMap +
  stat_density2d(data = data,
                 aes(Longitude, Latitude, fill = ..level..),
                 geom = "polygon", alpha = 0.7) +
  scale_fill_gradient(low = "white", high = "darkred") +
  facet_wrap(MobileOS~dow, nrow = 2) +
  ggtitle("Daily Usage of Mobile OS Accross LA")



```

Saturdays and Sundays in the North Hollywood area are busy for iOS





