---
title: "Explore 2015-2016 Dataset"
output: html_document
date: "11/26/2016"
---

**Load Add-on packages**
```{r message = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
```

**Data Preparation**

Import Data and transfer time-related variables to date/time objects with lubridate package
```{r}
LA311Call_1516 = read.csv("MyLA311_Service_Request_Data_2016.csv")

LA311Call_1516$CreatedDate = mdy_hms(LA311Call_1516$CreatedDate)
LA311Call_1516$UpdatedDate = mdy_hms(LA311Call_1516$UpdatedDate)

LA311Call_1516$Year = year(LA311Call_1516$CreatedDate) 
LA311Call_1516$Month = month(LA311Call_1516$CreatedDate, label = TRUE)
LA311Call_1516$DoW = wday(LA311Call_1516$CreatedDate, label = TRUE)
LA311Call_1516$Hour = hour(LA311Call_1516$CreatedDate)
```

**Request Type Breakdown**

General distribution of request type
```{r}
Request_Type = LA311Call_1516 %>%
  group_by(RequestType) %>%
  summarise(Count = n())

ggplot(Request_Type, aes(x = reorder(RequestType, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "lightblue", color = "NA") +
  xlab("Request Type") +
  ylab("Count") +
  scale_y_continuous(breaks = seq(0, 600000, 100000)) +
  coord_flip() +
  ggtitle("Distribution of Request Type") +
  theme_minimal()
```

Distribution of request type categorized by request source
```{r}
Request_Source = LA311Call_1516 %>%
  filter(RequestSource %in% c("Call","Mobile App")) %>%
  filter(!is.na(RequestType))

Request_Source$RequestType = factor(Request_Source$RequestType, level = c("Feedback", "Report Water Waste", "Multiple Streetlight Issue", "Homeless Encampment", "Other", "Single Streetlight Issue", "Dead Animal Removal", "Electronic Waste", "Illegal Dumping Pickup", "Metal/Household Appliances", "Graffiti Removal", "Bulky Items"))

ggplot(Request_Source, aes(x = RequestType, fill = RequestSource)) +
  geom_bar(position = "dodge", width = 0.8) +
  scale_fill_manual(values = c("lightblue", "pink")) +
  xlab("Request Type") +
  ylab("") +
  ggtitle("Distribution of Request Type Categorized by Request Source") +
  theme_minimal() +
  coord_flip() 
```

Percentage of requests by hour and request type
```{r}
TypebyHour = LA311Call_1516 %>% 
  filter(!is.na(ServiceDate)) %>% 
  group_by(Hour, RequestType) %>%
  summarise(NumberofCalls = n())
  
TypebyHour = TypebyHour %>%
  group_by(Hour) %>%
  mutate(Percentage = NumberofCalls/sum(NumberofCalls) * 100)

ggplot(TypebyHour, aes(x = factor(Hour), y = Percentage, fill = RequestType, group = RequestType)) +
  geom_area() +
  geom_line(position = "stack", size = 0.1) +
  xlab("Hour") +
  theme_minimal() +
  guides(fill = guide_legend(reverse = T))
```

Percentage of requests by day of week and request type
```{r}
TypebyWeekday = LA311Call_1516 %>% 
  filter(!is.na(ServiceDate)) %>%
  group_by(DoW, RequestType) %>%
  summarise(NumberofCalls = n())
  
TypebyWeekday = TypebyWeekday %>%
  group_by(DoW) %>%
  mutate(Percentage = NumberofCalls/sum(NumberofCalls) * 100)

ggplot(TypebyWeekday, aes(x = DoW, y = Percentage, fill = RequestType, group = RequestType)) +
  geom_area() +
  geom_line(position = "stack", size = 0.1) +
  theme(legend.position = "none") +
  xlab("Day of Week") +
  ylab("Percentage(%)") +
  theme_minimal() +
  guides(fill = guide_legend(reverse = T))
```

Percentage of requests by month and request type
```{r}
TypebyMonth = LA311Call_1516 %>%
  group_by(Month, RequestType) %>%
  summarise(NumberofCalls = n())
  
TypebyMonth = TypebyMonth %>%
  group_by(Month) %>%
  mutate(Percentage = NumberofCalls/sum(NumberofCalls) * 100)

ggplot(TypebyMonth, aes(x = Month, y = Percentage, fill = RequestType, group = RequestType)) +
  geom_area() +
  geom_line(position = "stack", size = 0.1) +
  ylab("Percentage (%)") +
  theme_minimal() +
  guides(fill = guide_legend(reverse = T))
```

**Volume Breakdown**

Change volume of requests over time
```{r}
Request_Volume = LA311Call_1516 %>% 
  group_by(Year, Month) %>% 
  summarise(Count=n()) 

ggplot(Request_Volume, aes(x = Month, y = Count)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_line(color = "pink", group = 1) +
  geom_point(color = "red") +
  facet_wrap(~Year, nrow=1) + 
  ggtitle("Change Volume of Requests over Time") + 
  xlab("Month") + 
  ylab("Number of Requests") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 30, hjust = 1))
```

Change volume of requests by mobile App over time
```{r}
RequestSource_Volume = LA311Call_1516 %>% 
  filter(RequestSource == "Mobile App") %>%
  group_by(Year, Month) %>% 
  summarise(Count=n()) 

ggplot(RequestSource_Volume, aes(x = Month, y = Count)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_line(color = "pink", group = 1) +
  geom_point(color = "red") +
  facet_wrap(~Year, nrow=1) + 
  ggtitle("Change Volume of Requests by Mobile App over Time") + 
  xlab("Month") + 
  ylab("Number of Requests") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 30, hjust = 1))
```

Change volume of requests by call over time
```{r}
RequestSource_Volume = LA311Call_1516 %>% 
  filter(RequestSource == "Call") %>%
  group_by(Year, Month) %>% 
  summarise(Count=n()) 

ggplot(RequestSource_Volume, aes(x = Month, y = Count)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_line(color = "pink", group = 1) +
  geom_point(color = "red") +
  facet_wrap(~Year, nrow=1) + 
  ggtitle("Change Volume of Requests by Call over Time") + 
  xlab("Month") + 
  ylab("Number of Requests") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 30, hjust = 1))
```

Change volume of requests from Android and iOS over time
```{r}
Mobile_OS = LA311Call_1516 %>%
  filter(MobileOS == "iOS" | MobileOS == "Android") %>%
  group_by(Year, Month, MobileOS) %>%
  summarise(Count=n())

ggplot(Mobile_OS, aes(x = Month,y = Count, fill = MobileOS)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Year, nrow=1) +
  xlab("Month") + 
  ylab("Number of Requests") +
  scale_fill_manual(values = c("lightblue", "pink")) +
  ggtitle("Android vs. iOS Overtime") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 30, hjust = 1))
```

**Efficiency Analysis**

Distribution of average process time
```{r}
ProcessTime = LA311Call_1516 %>%
  mutate(ProcessTime = as.numeric(UpdatedDate - CreatedDate)/60/60/24) %>%
  filter(ProcessTime > 0) %>%
  group_by(RequestType, PolicePrecinct) %>%
  summarise(ProcessTime_Avg = round(mean(ProcessTime), 1))

ggplot(ProcessTime, aes(y = PolicePrecinct, x = RequestType, fill = ProcessTime_Avg)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "darkred") +
  ggtitle("Distribution of Average Process Time") +
  theme(axis.title = element_text(angle=0, colour="darkred", size=14),
        plot.title = element_text(colour="darkred", size=18, face="bold"),
        legend.title=element_text(colour="darkred", size=10),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1)) +
  labs(fill = "ProcessTime\n(days)")
```

