---
title: "Explore 2011-2015 Dataset"
output: html_document
date: "11/26/2016"
---
**Load Add-on packages**
```{r message = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
```

**Data Preparation**

Import Data and transfer time-related variables to date/time objects with lubridate package

```{r}
LA311Call_1415 = read.csv("311_Call_Center_Tracking_Data.csv")
LA311Call_1415$Date = mdy(LA311Call_1415$Date) 
LA311Call_1415$Time = hms(LA311Call_1415$Time)
```

Add two variables DoW and Hour indicating the day of week and hour of the request respectively.
```{r}
LA311Call_1415$Month=month(LA311Call_1415$Date,label=TRUE)
LA311Call_1415$DoW = wday(LA311Call_1415$Date, label = TRUE)
LA311Call_1415$Hour = hour(LA311Call_1415$Time)
```

**Visualization of Peak Season / Month / Hour**

Heatmap over month and hour
```{r}
heatmap_data_1 = LA311Call_1415 %>%
  filter(Month != "NA") %>%
  group_by(Month,Hour) %>%
  summarise(Count = n())

ggplot(heatmap_data_1,aes(x = Month,y = factor(Hour),fill = Count))+
  geom_tile()+
  scale_fill_gradient(low = "white",high = "darkred")+
  theme_minimal()+
  ylab("Hour")+
  ggtitle("Heatmap of 311 Requests in Los Angeles")
```

Heatmap over day of week and hour
```{r}
heatmap_data_2 = LA311Call_1415 %>%
  filter(!is.na(DoW))%>%
  group_by(DoW, Hour)%>%
  summarise(count = n())

ggplot(heatmap_data_2, aes(x = DoW, y = factor(Hour), fill = count)) + 
  geom_tile()+
  scale_fill_gradient(low = "white", high = "darkred") + xlab("Day of Week") + ylab("Hour") +
  theme_minimal() +
  ggtitle("Heatmap of 311 Requests in Los Angeles")
```

**Percentage of Requests Handled vs Referred**

Add a new variable to indicate wheather the request was handled or referred to other department
```{r}
LA311Call_1415 = 
    transform(
        LA311Call_1415,
        Resolution =
            ifelse(Call.Resolution %in% c("Referred To 411",
                                           "Referred To County",
                                           "Referred To Other Governmental",
                                           "Referred To State",
                                           "Transfer (City)",
                                           "Transferred To 411",
                                           "Warm Transfer (City)",
                                           "Got Voicemail (City)"), 
                   'Referred',
                   ifelse(Call.Resolution %in% c("Call Resolution",
                                          "Escalate To Supervisor",
                                          "Escalated To Office of Finance",
                                          "Gave Caller Information",
                                          "Service Request Processed"),
                          'Handled',
                          'NA')))
unique(LA311Call_1415$Department.Abbreviation)
```

Create a pie chart to display the percentage of handled request and referred request
```{r}
Resolution = LA311Call_1415 %>%
  filter(Resolution != "NA") %>%
  group_by(Resolution)%>%
  summarise(Count=n()) %>%
  mutate(Percent=Count*100/sum(Count))

ggplot(Resolution,aes(x = factor(1), y = Percent,fill = Resolution)) +
  geom_bar(stat = "identity", width = 1)+
  ylab("Percentage of Total")+
  ggtitle("Percentage of Handled vs. Referred")+
  theme_minimal() +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("lightblue", "pink")) +
  xlab("") +
  ylab("")
```

**Performance of Different Departments**

Top 10 departments with highest referred rate
```{r}
Department_Resolution = LA311Call_1415 %>%
  filter(!is.na(Resolution) & (Department.Abbreviation != ""))%>%
  group_by(Department.Abbreviation) %>%
  summarise(Referred_Percent = sum(Resolution == "Referred")/n(), 
            Handled_Percent = sum(Resolution == "Handled")/n())

Department_Resolution_Highest = subset(Department_Resolution, rank(Referred_Percent) > 51)

ggplot(Department_Resolution_Highest, 
       aes(x = reorder(Department.Abbreviation, -Referred_Percent), 
           y = Referred_Percent * 100)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  xlab("Department") +
  ylab("Percentage of Referred Request (%)") +
  ggtitle("Top 10 Departments with Highest Referred Rate") +
  theme_minimal() +
  coord_flip()
```

Top 10 departments with lowest referred rate
```{r}
Department_Resolution_Lowest = subset(Department_Resolution, rank(Referred_Percent) < 11)

ggplot(Department_Resolution_Lowest, 
       aes(x = reorder(Department.Abbreviation, -Referred_Percent), 
           y = Referred_Percent * 100)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  xlab("Department") +
  ylab("Percentage of Referred Request (%)") +
  ggtitle("Top 10 Departments with Lowest Referred Rate") +
  theme_minimal() +
  coord_flip()
```

Top 10 departments receiving most requests
```{r}
Department_Count = LA311Call_1415 %>%
  group_by(Department.Abbreviation) %>%
  summarise(Count = n()) %>%
  filter(Department.Abbreviation != "") %>%
  filter(rank(Count) > 51)

ggplot(Department_Count, aes(x = Count, 
                             y = reorder(Department.Abbreviation,Count))) +
  geom_point(size = 2) +
  xlab("Number of Requests Received") +
  ylab("Department") +
  ggtitle("Top 10 Departments Receiving Most Requests") +
  theme_minimal()
```

**Geographic Visualization**

Distribution of requests displayed on map
```{r}
zipcode=read.csv("zipcode.csv")

map_data = merge(LA311Call_1415, 
                 zipcode,
                 by.x = "Zip.Code",
                 by.y = "ZIP",
                 all.x = TRUE)

map_data_filtered = filter(map_data, !is.na(LNG) | !is.na(LAT) )

library(ggmap)
LA = qmap("Los Angeles", zoom = 11, maptype = "roadmap", legend = "topleft")

LA+
  stat_density2d(data = map_data_filtered,              
                 aes(x = LNG ,y = LAT, fill=..level..),  
                 geom = "polygon") +
  scale_fill_gradient(low = "gray", high = "darkred")
```

