---
title: "Final project R code"
author: "DSO 545: Statistical Computing and Data Visualization"
date: "Fall 2016"
output: pdf_document
---


1. Distribution of Department in Call Center Dataset

```{r,warning=FALSE}
callcenter=read.csv("callcenter.csv")
servicerequest=read.csv("servicerequest.csv")
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggmap)

call_dep=callcenter %>%
  group_by(Department.Abbreviation) %>%
  summarise(count=n())
ggplot(call_dep,aes(x=reorder(Department.Abbreviation,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Distribution of Departments in Call Center")+
  theme_bw()

```

2. Distribution of Department in Service Request Dataset
```{r, warning=FALSE}
service_dep=servicerequest %>%
  filter(Owner != "")%>%
  group_by(Owner) %>%
  summarise(count=n())

ggplot(service_dep,aes(x=reorder(Owner,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Distribution of Department in Service Request")+
  theme_bw()

```

3. Distribution of Request Source in Service Request Dataset
```{r, warning=FALSE}
service_sor = servicerequest %>%
  group_by(RequestSource) %>%
  summarise(count=n())
ggplot(service_sor,aes(x=reorder(RequestSource,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Request Source")+
  ggtitle("Distribution of Request Source")+
  theme_bw()

```

4.Request Source in Each Department
```{r, warning=FALSE}
service= servicerequest %>%
  filter(Owner!="")
ggplot(service,aes(x=RequestSource))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Request Source in Each Department")+
  theme_bw()+
  facet_wrap(~Owner)
```

5. Request Source in Each Request Type 
```{r, warning=FALSE}
ggplot(servicerequest,aes(x=RequestType,fill=RequestSource))+
  geom_bar()+
  coord_flip()+
  ylab("Count")+
  xlab("Request")+
  ggtitle("Request Source in Each Request")+
  theme_bw()

```

6. Distribution of Call Resolution
```{r, warning=FALSE}
call_resolu = callcenter%>%
  filter(Call.Resolution !="N/A") %>%
  filter(Call.Resolution!="Call Resolution")%>%
  group_by(Call.Resolution) %>%
  summarise(count=n())

ggplot(call_resolu,aes(x=reorder(Call.Resolution,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Call Resolution")+
  ggtitle("Distribution of Call Resolution")+
  theme_bw()

```

7. Service Request Density Map
```{r, warning=FALSE}
library(ggmap)
LA = qmap("Los Angeles",zoom = 10)
LA+ stat_density2d(data=servicerequest,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map")

```

8. Call Center Heatmap
```{r, warning=FALSE}
callcenter$Date=mdy(callcenter$Date)
callcenter$dow = wday(callcenter$Date,label=T)
callcenter$Time=hms(callcenter$Time)
callcenter$hour =hour(callcenter$Time)
heatmap=callcenter %>%
  filter(!is.na(dow))%>%
  group_by(dow,hour) %>%
  summarise(count=n())


ggplot(heatmap,aes(x=dow,y=hour,fill=count))+
  geom_tile(alpha=0.7)+
  scale_fill_gradient(low="white",high="darkred")+
  xlab("Weekday")+
  ylab("Hour") +
  ggtitle("Call Center Heatmap")

```

9. Service Request Map by Weekday
```{r, warning=FALSE}
servicerequest$ServiceDate=mdy_hms(servicerequest$ServiceDate)
servicerequest$dow = wday(servicerequest$ServiceDate,label=T)
servicerequest$month = month(servicerequest$ServiceDate,label=T)

dowdata = servicerequest %>%
  filter(!is.na(dow))
LA+ stat_density2d(data=dowdata,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map by Weekday")+
  facet_wrap(~dow)
```

10. Service Qequest Map by Month
```{r, warning=FALSE}
monthdata = servicerequest %>%
  filter(!is.na(month))
  LA+ stat_density2d(data=monthdata,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map by Month")+
  facet_wrap(~month)

```

11. Change Volume of Reuqest Over Month
```{r, warning=FALSE}
servicerequest %>%
  filter(!is.na(month)) %>%
  ggplot(aes(x=month))+
  geom_bar(fill="lightblue",color="lightblue")+
  xlab("Month")+
  ylab("Request")+
  ggtitle("Change Volume of Request Over Month")

```

12. Change Volume of Request Over Weekday
```{r, warning=FALSE}
servicerequest %>%
  filter(!is.na(dow)) %>%
  ggplot(aes(x=dow))+
  geom_bar(fill="lightblue",color="lightblue")+
  xlab("Weekday")+
  ylab("Request")+
  ggtitle("Change Volume of Request Over Weekday")
```

13. Distribution of Request Type
```{r, warning=FALSE}
service_type=servicerequest%>%
  group_by(RequestType) %>%
  summarise(count=n())
ggplot(service_type,aes(x=reorder(RequestType,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("RequestType")+
  ylab("Count")+
  ggtitle("Distribution of Request Type")
```

14. Distribution of Call resolution Type
```{r, warning=FALSE}
callcenter=callcenter %>%
  mutate(Call.Res.Class=ifelse(Call.Resolution %in% c("Transfer (City)","Warm Transfer (City)","Transferred to 411"), "Transfer",
                             ifelse(Call.Resolution =="Service Request Processed", "Service Request",
                                    ifelse(Call.Resolution =="Gave Caller Information", "Gave Info", 
                                           ifelse(Call.Resolution %in% c("Referred To County", "Referred To State", "Referred To Other Governmental", "Referred To 411"), "Refer", 
                                                  "Others")))))
call_resolustion = callcenter%>%
  group_by(Call.Res.Class)%>%
  summarise(count=n())
ggplot(call_resolustion,aes(x=reorder(Call.Res.Class,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("Call Resolution Type")+
  ylab("Count")+
  ggtitle("Distribution of Call Resolution Type")+
  theme_bw()
```

15. Call Resolution Type for Each Department
```{r, warning=FALSE}
ggplot(callcenter,aes(x=Department.Abbreviation,fill=Call.Res.Class))+
  geom_bar()+
  coord_flip()+
  xlab("Department.Name")+
  ylab("Count")+
  ggtitle("Call Resolution Type for Each Department")+
  theme_bw()
```

16. Distribution of Top 15 AssignTo
```{r, warning=FALSE}
service_at=servicerequest %>%
  filter(AssignTo !="")%>%
  group_by(AssignTo) %>%
  summarise(count=n()) %>%
  arrange(-count)%>%
  slice(1:15)
ggplot(service_at,aes(x=reorder(AssignTo,count),y=count))+
  geom_bar(stat="identity",color="lightblue",fill="lightblue")+
  coord_flip()+
  xlab("AssignTo")+
  ylab("Count")+
  ggtitle("Distribution of Top 15 AssignTo")

```

17. AssignTo for Each Request Type
```{r, warning=FALSE}
service_ta=servicerequest%>%
  filter(AssignTo %in% c("NC","EV","SLA","WV","WLA","SLABT",
                         "HB","GAP","GAPBH","CCAC","WVA",
                         "HBT","KYCC","LACC",
                         "NEGB_ELA"))
ggplot(service_ta,aes(x=AssignTo))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  facet_wrap(~RequestType)+
  xlab("AssignTo")+
  ylab("Count")+
  ggtitle("AssignTo for Each Reuqest Type")
```

18. Distribution of APC
```{r, warning=FALSE}
service_apc=servicerequest %>%
  filter(APC !="") %>%
  group_by(APC) %>%
  summarise(count=n())
ggplot(service_apc,aes(x=reorder(APC,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("APC")+
  ylab("Count")+
  ggtitle("Distribution of APC")+
  theme_bw()

```

19. Request Type Distribution for Each APC
```{r, warning=FALSE}
servicerequest %>%
  filter(APC !="") %>%
  ggplot(aes(x=RequestType))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  facet_wrap(~APC)+
  theme_bw()+
  xlab("Request Type")+
  ylab("Count")+
  ggtitle("Request Type Distribution for Each APC")

```

20. Distribution of Average Process Time for Each Request Type
```{r, warning=FALSE}
servicerequest$CreatedDate=mdy_hms(servicerequest$CreatedDate)
servicerequest$UpdatedDate=mdy_hms(servicerequest$UpdatedDate)
servicerequest=servicerequest %>%
  mutate(process_time=(UpdatedDate-CreatedDate)/3600)
service_time=servicerequest %>%
  group_by(RequestType) %>%
  summarise(avgtime=mean(process_time))
ggplot(service_time,aes(x=reorder(RequestType,avgtime),y=avgtime))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("Request Type")+
  ylab("Average Process Time(Hour)")+
  ggtitle("Distribution of Average Process Time for Each Request Type")+
  theme_bw()
```

21. Service Request VS. Process Time
```{r, warning=FALSE}
ggplot(servicerequest,aes(x=RequestType,y=process_time))+
  geom_point(position=position_jitter(height=0,width=0.4),alpha=0.5)+
  xlab("Request Type")+
  ylab("Process Time")+
  ggtitle("Request Type VS. Process Time")+
  coord_flip()+
  theme_bw()
```

22. Distribution of Average Process Time for Each AssignTo(Top 15)
```{r, warning=FALSE}
service_time2=servicerequest %>%
  group_by(AssignTo) %>%
  summarise(avgtime=mean(process_time)) %>%
  arrange(-avgtime) %>%
  slice(1:15)

ggplot(service_time2,aes(x=reorder(AssignTo,avgtime),y=avgtime))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("AssignTo")+
  ylab("Average Process Time(Hour)")+
  ggtitle("Distribution of Average Process Time for Each AssignTo(Top 15)")+
  theme_bw()
```

23. Request Type for Each Council District
```{r, warning=FALSE}
servicerequest%>%
  filter(!is.na(CD))%>%
  ggplot(aes(x=RequestType))+
  geom_bar(fill="lightblue",color="lightblue")+
  facet_wrap(~CD)+
  xlab("Request Type")+
  ylab("Count")+
  ggtitle("Request Type for Each Council District")+
  coord_flip()+
  theme_bw()
```

24.Distribution of Requests for Council Districts
```{r, warning=FALSE}
servicerequest %>%
  filter(!is.na(CD)) %>%
  ggplot(aes(x=as.factor(CD)))+
  geom_bar(fill="lightblue",color="lightblue")+
  xlab("Council Districts")+
  ylab("Count")+
  ggtitle("Distribution of Requests for Council Districts")+
  theme_bw()

```

25.Distribution of Requests for Median Household income and Bachelor Degree
```{r, warning=FALSE}
# Median Household Income
income=read.csv("Income.csv")
# Educational Attainment for the Population 25 Years and Over
education=read.csv("education.csv")
# population by Zip Code
population=read.csv("population.csv")

education<-education %>%
  filter(!is.na(BachelorAbove)) %>%
  select(ZipCode,BachelorAbove)

# BachelorAbove is the % of people who have bachelor's degree or above

subset<-servicerequest %>%
    filter(!is.na(ZipCode)) %>%
    group_by(ZipCode,RequestType) %>%
    summarise(Request=n())

request_population <- merge(population,subset,by="ZipCode")
request_population<-mutate(request_population,request_10000=Request/Population*10000)

income_request <- merge(income,request_population,by="ZipCode")

education_request <- merge(education,request_population,by="ZipCode")

income_request<-income_request %>%
  filter(request_10000<2000)

education_request<-education_request %>%
  filter(request_10000<2000)

ggplot(income_request,aes(x=Income,y=request_10000))+
  geom_point()+
  facet_wrap(~RequestType,scales="free")+
  xlab("Median Household Income")+
  ylab("Requests per 10,000 pop")

ggplot(education_request,aes(x=BachelorAbove,y=request_10000))+
  geom_point()+
  facet_wrap(~RequestType,scales="free")+
  xlab("Bachelor Degree or Above %")+
  ylab("Requests per 10,000 pop")

```

26.Service Request Map
```{r,warning=FALSE}
LA = qmap("Los Angeles",zoom = 11)
LA+ stat_density2d(data=servicerequest,aes(x=Longitude,y=Latitude,
                                           fill=..level..,alpha=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="gray",high="darkred")+
  facet_wrap(~RequestType,nrow=3)+
  ggtitle("Service Request Map")
```

27.Change Volume of Request Type Over Time
```{r,warning=FALSE}
servicerequest$ServiceDate=mdy_hms(servicerequest$ServiceDate)
servicerequest$dow = wday(servicerequest$ServiceDate,label=T)
data=servicerequest %>%
  filter(!is.na(dow)) %>%
  droplevels()
subdata1=data %>%
  filter(RequestType=="Bulky Items"|
           RequestType=="Electronic Waste"|
           RequestType=="Graffiti Removal"|
           RequestType=="Illegal Dumping Pickup"|
           RequestType=="Metal/Household Appliances"|
           RequestType=="Single Streetlight Issue") %>%
  droplevels()
ggplot(subdata1,aes(x=dow,fill=RequestType))+
  geom_bar(position="dodge",width=2)+
  facet_wrap(~RequestType,nrow=2)+
  xlab("Weekday")+
  ylab("Request")+
  ggtitle("Change Volume of Request Type Over Time")
```

28.Change Volume of Input Type Over Time
```{r,warning=FALSE}
subdata2=data %>%
  filter(RequestSource=="Call"|
           RequestSource=="Driver Self Report"|
           RequestSource=="Email"|
           RequestSource=="Mobile App") %>%
  droplevels()
  ggplot(subdata2,aes(x=dow,fill=RequestSource))+
  geom_bar(position="dodge",width=3)+
  facet_wrap(~RequestSource,nrow=3)+
  xlab("Weekday")+
  ylab("Request")+
  ggtitle("Change Volume of Input Type Over Time")
```

29.Change Volume of Request Type Over Time
```{r,warning=FALSE}
servicerequest$month = month(servicerequest$ServiceDate,label=T)
data2=subset(servicerequest,!is.na(month))
subdata3=data2 %>%
  filter(RequestType=="Bulky Items"|
           RequestType=="Electronic Waste"|
           RequestType=="Graffiti Removal"|
           RequestType=="Illegal Dumping Pickup"|
           RequestType=="Metal/Household Appliances"|
           RequestType=="Single Streetlight Issue") %>%
  droplevels()
  ggplot(subdata3,aes(x=month,fill=RequestType))+
  geom_bar(position="dodge",width=2)+
  facet_wrap(~RequestType,nrow=3)+
  xlab("Month")+
  ylab("Request")+
  ggtitle("Change Volume of Request Type Over Time")
```

30.Change Volume of Input Type Over Time
```{r,warning=FALSE}
  subdata2$month = month(subdata2$ServiceDate,label=T)
  ggplot(subdata2,aes(x=month,fill=RequestSource))+
  geom_bar(position="dodge",width=3)+
  facet_wrap(~RequestSource,nrow=3)+
  xlab("Month")+
  ylab("Request")+
  ggtitle("Change Volume of Input Type Over Time")
```

31.Number of Service by Day
```{r,warning=FALSE}
 callcenter$Date <- mdy(callcenter$Date)
  callcenter$Time=hms(callcenter$Time)
  callcenter$hour=hour(callcenter$Time)
  callcenter$day=wday(callcenter$Date,label = T, abbr = F)
  summary(callcenter$day)
  lev=levels(callcenter$day)
  lev=lev[c(2:7,1)]
  callcenter$day=factor(callcenter$day,levels=lev)
  callcenter %>%
    filter(!is.na(day)) %>%
    ggplot(aes(x=day))+
    geom_bar(fill="lightblue")+
    xlab("Weekday")+
    ylab("Service")+
    ggtitle("Number of Service by Day")
```

32.Change Volume of Request Type Over Weekday
```{r,warning=FALSE}
servicerequest$ServiceDate=mdy_hms(servicerequest$ServiceDate)
servicerequest$dow = wday(servicerequest$ServiceDate,label=T)
ggplot(servicerequest,aes(x=dow,colour=RequestType))+
  geom_line(stat="count",aes(group=RequestType),size=1)+
  xlab("Weekday")+
  ylab("Request")+
  ggtitle("Change Volume of Request Type Over Weekday")+
  theme_bw()
```

33.Change Volume of Request Type Over Month
```{r,warning=FALSE}
servicerequest$month = month(servicerequest$ServiceDate,label=T)
ggplot(servicerequest,aes(x=month,colour=RequestType))+
  geom_line(stat="count",aes(group=RequestType),size=1)+
  xlab("Month")+
  ylab("Request")+
  ggtitle("Change Volume of Request Type Over Month")+
  theme_bw()
```

34.Distribution of Process Time for RequestType
```{r,warning=FALSE}
ggplot(servicerequest,aes(x=RequestType,y=process_time))+
  geom_boxplot(color="blue")+
  coord_flip()+
  ggtitle("Distribution of Process Time for RequestType")+
  theme_bw()
```
