---
title: "R Code"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Distribution of Department in Call Center Dataset

```{r}
callcenter=read.csv("callcenter.csv")
servicerequest=read.csv("servicerequest.csv")
library(ggplot2)
library(dplyr)
library(lubridate)

call_dep=callcenter %>%
  group_by(Department.Abbreviation) %>%
  summarise(count=n())
ggplot(call_dep,aes(x=reorder(Department.Abbreviation,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Distribution of Departments in Call Center")+
  theme_bw()

```


2. Distribution of Department in Service Request Dataset
```{r}
service_dep=servicerequest %>%
  filter(Owner != "")%>%
  group_by(Owner) %>%
  summarise(count=n())

ggplot(service_dep,aes(x=reorder(Owner,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Distribution of Department in Service Request")+
  theme_bw()

```

3. Distribution of Request Source in Service Request Dataset
```{r}
service_sor = servicerequest %>%
  group_by(RequestSource) %>%
  summarise(count=n())
ggplot(service_sor,aes(x=reorder(RequestSource,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Request Source")+
  ggtitle("Distribution of Request Source")+
  theme_bw()

```

4.Request Source in Each Department
```{r}
service= servicerequest %>%
  filter(Owner!="")
ggplot(service,aes(x=RequestSource))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Department")+
  ggtitle("Request Source in Each Department")+
  theme_bw()+
  facet_wrap(~Owner)
```

5. Request Source in Each Request Type 
```{r}
ggplot(servicerequest,aes(x=RequestType,fill=RequestSource))+
  geom_bar()+
  coord_flip()+
  ylab("Count")+
  xlab("Request")+
  ggtitle("Request Source in Each Request")+
  theme_bw()

```

6. Distribution of Call Resolution
```{r}
call_resolu = callcenter%>%
  filter(Call.Resolution !="N/A") %>%
  filter(Call.Resolution!="Call Resolution")%>%
  group_by(Call.Resolution) %>%
  summarise(count=n())

ggplot(call_resolu,aes(x=reorder(Call.Resolution,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  ylab("Count")+
  xlab("Call Resolution")+
  ggtitle("Distribution of Call Resolution")+
  theme_bw()

```

7. Service Request Density Map
```{r}

library(ggmap)
LA = qmap("Los Angeles",zoom = 10)
LA+ stat_density2d(data=servicerequest,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map")

```

8. Call Center Heatmap
```{r}
callcenter$Date=mdy(callcenter$Date)
callcenter$dow = wday(callcenter$Date,label=T)
callcenter$Time=hms(callcenter$Time)
callcenter$hour =hour(callcenter$Time)
heatmap=callcenter %>%
  filter(!is.na(dow))%>%
  group_by(dow,hour) %>%
  summarise(count=n())


ggplot(heatmap,aes(x=dow,y=hour,fill=count))+
  geom_tile(alpha=0.7)+
  scale_fill_gradient(low="white",high="darkred")+
  xlab("Weekday")+
  ylab("Hour") +
  ggtitle("Call Center Heatmap")+
  theme_bw()

```

9. Service Request Map by Weekday
```{r}
servicerequest$ServiceDate=mdy_hms(servicerequest$ServiceDate)
servicerequest$dow = wday(servicerequest$ServiceDate,label=T)
servicerequest$month = month(servicerequest$ServiceDate,label=T)

dowdata = servicerequest %>%
  filter(!is.na(dow))
LA+ stat_density2d(data=dowdata,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map by Weekday")+
  facet_wrap(~dow)
```

10. Service Qequest Map by Month
```{r}
monthdata = servicerequest %>%
  filter(!is.na(month))
  LA+ stat_density2d(data=monthdata,aes(x=Longitude,y=Latitude,
                                           fill=..level..),
                   alpha=0.5,geom="polygon")+
  scale_fill_gradient(low="grey",high="darkred")+
  ggtitle("Service Request Map by Month")+
  facet_wrap(~month)

```

11. Change Volume of Reuqest Over Month
```{r}
service_month=servicerequest %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(count=n())
ggplot(service_month,aes(x=month,y=count,group=1))+
  geom_line(color="lightblue")+
  xlab("Month")+
  ylab("Request")+
  ggtitle("Change Volume of Request Over Month")+
  theme_bw()+
  geom_point()

```

12. Change Volume of Request Over Weekday
```{r}
service_weekday=servicerequest %>%
  filter(!is.na(dow)) %>%
  group_by(dow) %>%
  summarise(count=n())
ggplot(service_weekday,aes(x=dow,y=count,group=1))+
  geom_line(color="lightblue")+
  xlab("Weekday")+
  ylab("Request")+
  ggtitle("Change Volume of Request Over Weekday")+
  geom_point()+
  theme_bw()
```

13. Distribution of Request Type
```{r}
service_type=servicerequest%>%
  group_by(RequestType) %>%
  summarise(count=n())
ggplot(service_type,aes(x=reorder(RequestType,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("RequestType")+
  ylab("Count")+
  ggtitle("Distribution of Request Type")+
  theme_bw()
```

14. Distribution of Call resolution Type
```{r}
callcenter=callcenter %>%
  mutate(Call.Res.Class=ifelse(Call.Resolution %in% c("Transfer (City)","Warm Transfer (City)","Transferred to 411"), "Transfer",
                             ifelse(Call.Resolution =="Service Request Processed", "Service Request",
                                    ifelse(Call.Resolution =="Gave Caller Information", "Gave Info", 
                                           ifelse(Call.Resolution %in% c("Referred To County", "Referred To State", "Referred To Other Governmental", "Referred To 411"), "Refer", 
                                                  "Others")))))
call_resolustion = callcenter%>%
  group_by(Call.Res.Class)%>%
  summarise(count=n())
ggplot(call_resolustion,aes(x=reorder(Call.Res.Class,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("Call Resolution Type")+
  ylab("Count")+
  ggtitle("Distribution of Call Resolution Type")+
  theme_bw()
```

15. Call Resolution Type for Each Department
```{r}
ggplot(callcenter,aes(x=Department.Abbreviation,fill=Call.Res.Class))+
  geom_bar()+
  coord_flip()+
  xlab("Department.Name")+
  ylab("Count")+
  ggtitle("Call Resolution Type for Each Department")+
  theme_bw()
```

16. Distribution of Top 15 AssignTo
```{r}
service_at=servicerequest %>%
  filter(AssignTo !="")%>%
  group_by(AssignTo) %>%
  summarise(count=n()) %>%
  arrange(-count)%>%
  slice(1:15)
ggplot(service_at,aes(x=reorder(AssignTo,count),y=count))+
  geom_bar(stat="identity",color="lightblue",fill="lightblue")+
  coord_flip()+
  xlab("AssignTo")+
  ylab("Count")+
  ggtitle("Distribution of Top 15 AssignTo")

```

17. AssignTo for Each Request Type
```{r}
service_ta=servicerequest%>%
  filter(AssignTo %in% c("NC","EV","SLA","WV","WLA","SLABT",
                         "HB","GAP","GAPBH","CCAC","WVA",
                         "HBT","KYCC","LACC",
                         "NEGB_ELA"))
ggplot(service_ta,aes(x=AssignTo))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  facet_wrap(~RequestType)+
  xlab("AssignTo")+
  ylab("Count")+
  ggtitle("AssignTo for Each Reuqest Type")
```

18. Distribution of APC
```{r}
service_apc=servicerequest %>%
  filter(APC !="") %>%
  group_by(APC) %>%
  summarise(count=n())
ggplot(service_apc,aes(x=reorder(APC,count),y=count))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("APC")+
  ylab("Count")+
  ggtitle("Distribution of APC")+
  theme_bw()

```

19. Request Type Distribution for Each APC
```{r}
servicerequest %>%
  filter(APC !="") %>%
  ggplot(aes(x=RequestType))+
  geom_bar(fill="lightblue",color="lightblue")+
  coord_flip()+
  facet_wrap(~APC)+
  theme_bw()+
  xlab("Request Type")+
  ylab("Count")+
  ggtitle("Request Type Distribution for Each APC")

```

20. Distribution of Average Process Time for Each Request Type
```{r}
servicerequest$CreatedDate=mdy_hms(servicerequest$CreatedDate)
servicerequest$UpdatedDate=mdy_hms(servicerequest$UpdatedDate)
servicerequest=servicerequest %>%
  mutate(process_time=(UpdatedDate-CreatedDate)/3600)
service_time=servicerequest %>%
  group_by(RequestType) %>%
  summarise(avgtime=mean(process_time))
ggplot(service_time,aes(x=reorder(RequestType,avgtime),y=avgtime))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("Request Type")+
  ylab("Average Process Time(Hour)")+
  ggtitle("Distribution of Average Process Time for Each Request Type")+
  theme_bw()


```

21. Service Request VS. Process Time
```{r}
ggplot(servicerequest,aes(x=RequestType,y=process_time))+
  geom_point(position=position_jitter(height=0,width=0.4),alpha=0.5)+
  xlab("Request Type")+
  ylab("Process Time")+
  ggtitle("Request Type VS. Process Time")+
  coord_flip()+
  theme_bw()
```

22. Distribution of Average Process Time for Each Department
```{r}
service_time2=servicerequest %>%
  group_by(Owner) %>%
  summarise(avgtime=mean(process_time)) %>%
  arrange(-avgtime)

ggplot(service_time2,aes(x=reorder(Owner,avgtime),y=avgtime))+
  geom_bar(stat="identity",fill="lightblue",color="lightblue")+
  coord_flip()+
  xlab("Department")+
  ylab("Average Process Time(Hour)")+
  ggtitle("Distribution of Average Process Time for Each Department")+
  theme_bw()
```

23. Request Type for Each Council District
```{r}
servicerequest%>%
  filter(!is.na(CD))%>%
  ggplot(aes(x=RequestType))+
  geom_bar(fill="lightblue",color="lightblue")+
  facet_wrap(~CD)+
  xlab("Request Type")+
  ylab("Count")+
  ggtitle("Request Type for Each Council District")+
  coord_flip()+
  theme_bw()
```

24.Distribution of Requests for Council Districts
```{r}
servicerequest %>%
  filter(!is.na(CD)) %>%
  ggplot(aes(x=as.factor(CD)))+
  geom_bar(fill="lightblue",color="lightblue")+
  xlab("Council Districts")+
  ylab("Count")+
  ggtitle("Distribution of Requests for Council Districts")+
  theme_bw()

```

25. Sankey Diagram: Owner-RequestType
```{r}
library(googleVis)
service_sankey = servicerequest %>%
  group_by(Owner,RequestType) %>%
  summarise(count1=n())

plot(gvisSankey(service_sankey,from="Owner",to="RequestType",weight="count1",
                options=list(height=800, width=850,
                             sankey="{
                             link:{color:{fill: 'lightgray', fillOpacity: 0.7}},
                             node:{nodePadding: 5, label:{fontSize: 12}, interactivity: true, width: 20},
                             }")
)
)
```

26. Sankey Diagram: RequestType-AssignTo
```{r}
service_sankey2 = servicerequest %>%
  group_by(RequestType,AssignTo) %>%
  summarise(count2=n())

plot(gvisSankey(service_sankey2,from="RequestType",to="AssignTo",weight="count2",
                options=list(height=800, width=850,
                             sankey="{
                             link:{color:{fill: 'lightgray', fillOpacity: 0.7}},
                             node:{nodePadding: 5, label:{fontSize: 12}, interactivity: true, width: 20},
                             }")
)
)
```

27. change volume of request source by weekday
```{r}
ggplot(servicerequest,aes(x=dow,color=RequestSource))+
  geom_line(stat="count",aes(group=RequestSource))+
  xlab("Weekday")+
  ylab("Count")+
  ggtitle("Change Volume of Request Source by Weekday")+
  theme_bw()
```

28.Change Volume of Request Source by Month
```{r}
ggplot(servicerequest,aes(x=month,color=RequestSource))+
  geom_line(stat="count",aes(group=RequestSource))+
  xlab("Month")+
  ylab("Count")+
  ggtitle("Change Volume of Request Source by Month")+
  theme_bw()
```

